\thetitle

### 2013-04-21 tfpie/tfpie.tex
More on world.

>>>>

The key difference between these two programs is that the functional
program uses the {\tt 2htdp/universe} library, which provides a {\tt
  big-bang} form that consumes the initial state of the world and has
a declarative form of associating event-handler functions.  The
object-oriented program uses an alternative library developed for the
class language: {\tt class/universe}.  It also provides a {\tt
  big-bang} form but it consumes a single argument, the initial state
of the world represented as an object.  Event handlers are just
methods of this object.

The program on the left is the first program of the first semester,
while the one on the right is the first program of the second
semester.  Our approach is able to make the conceptual connection
between functional and object-oriented programming quite clear while
appealing the familiar event-driven interactive programs developed
throughout the year.

The move to object-oriented style immediately and naturally leads to
design principles that are enabled by organizing programs around
objects.  For example, the state-pattern becomes useful almost
immediately.  The programs in figure~\ref{fig:world} animate a rocket
(rendered as a circle in this example) taking off.  An illustrative
follow-up exercise is to animate a rocket that \emph{lands}.  The
natural design is to have two variants for states of the rocket: one
for descending rockets, one for landed rockets.  While in the
functional approach it is easy to use the state-pattern for the
\emph{data} representing a rocket, it is more difficult to have states
of \emph{behavior}.  Of course in the object approach, states of
behavior are just as natural as data.  Therefore it is straightforward
to design programs with easy to observe invariants such as ``a landed
rocket never changes position.''  In the functional approach, even such
simple properties are more involved to establish.
